Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-05-27 3dpas_intro.YS.tex
new files

\usepackage{graphicx} \usepackage{url} \usepackage{color} \usepackage{enumerate}
\usepackage{longtable} %\usepackage{authblk} % Is used internally within the CPE
Class \usepackage[firstpage]{draftwatermark}

@@

%\item Compute resource selection based on data location, processing profile
and/or network capacity

@@

%\item Collection and processing of monitoring information to steer grid
enactment.

@@

%\item Infrastructure monitoring for the execution of workflows to % manage
failure
%\item the application is sequence alignment (split data (this is where there is
dynamic decision making) to execute the alignment in parallel)
%\item DTI Imaging: two cases are interesting: DTI atlas (split, run, merge,
split, merge - reduction from 10000 tasks to 1 output); PCA for classification
of patients/control

@@

%The application performs analysis of DNA sequencing data for various %goals in
life science research, e.g., Mutation screening, Virus %discovery, Genome-wide
Associations (GWA), Linkage analysis, %Comparison of bacterial genomes, microRNA
expression and Exome %sequencing. The data analysis is composed of various
methods that are %implemented by a bioinformatician, whereas the data is
typically owned %by the life scientist.

@@

%bioinformatician, excluding steps related to data acquisition %(transformation
of raw images into sequences of amino acids) and %posterior biostatistics and
epidemiological analysis.
%The data analysis is implemented as a pipeline of generic and %customized
components, mostly third-party public software tools. The %pipeline may differ
for each specific study, depending on the data %acquisition and the research
question. Typically it includes data %preparation (file conversion, reformatting
and filtering); alignment %(comparison) with some reference database as the
human genome; %post-processing (output conversion, reformatting, statistical

@@

%computing-intensive step, which became prohibitive for sequential %execution in
next generation sequence experiments. All data is stored %in files that are
often completely read in memory for fast %processing. Compressed formats are
usually adopted. Although the files %are large, the data can be easily
decomposed because the processing is %done on each sequence (`string') at a
time.

@@

%Grid storage: Storage Elements of the Dutch e-Science Grid %(www.biggrid.nl)

@@

%Between data server and grid storage: public network (XXX); lightpath %(near
future)
%Between grid storage and worker nodes: public (XXX) or site (XXX) %network

@@

%Local Computing: clusters and servers at the AMC and workstations %(laptops,
desktops) anywhere Grid Computing: Dutch Grid %(www.biggrid.nl), part of EGI
%Originally all steps were performed in some local server or on the %user's
workstation. The visualization is still performed in the user's %(windows)
workstation, but all the other steps are currently performed %on the Dutch grid
infrastructure. The pipeline is described as a grid %workflow using the GWENDIA
language %(\url{http://dx.doi.org/10.1145/1645164.1645171}) and enacted on the

@@

%PS: although only the alignment step is computation-intensive, it is %more
practical to perform all the others on the grid to minimize data %transport
to/from the user's workstation, which is normally on low %speed networks. That
is, the computation is statically and manually %taken to the data in this
case. This can lead to sub-optimal resource %usage.

@@

%Note: data is not really `dynamic' in this case -- the sequences are %acquired
once and analyzed many times. However the execution of this %application on a
distributed infrastructure has dynamic aspects when %optimized resource usage is
considered:

@@

%DNA sequences (strings) stored in open formats (FASTA, SFF, SAM, BAM) %Bundles
of results from alignment tools (custom format, binary?)

@@

%input data: Sequencing Data in the csFasta format, normally 25-35~GB; %Quality
files in the .qual format, normally 50-80~GB; Reference DB in %the Fasta BS
format: 3.2~GB (human genome) 140~MB (one chromosome)
%Intermediate files: Reference BWA index, normally 4.5GB (human genome) %240~MB
(one chromosome); Sequencing Data in the FastQ format %(fastq.gz), normally
20-30~GB
%Output: Results in .sai format (direct output of BWA), normally %2-3~GB;
Results in .sam format, normally 55-75~GB; Results in .bam %format, normally
20-30~GB

@@

%From a local file. The legacy code is wrapped to stage the inputs and %outputs
in/out the worker node.
%All data is stored in files that are often completely read in memory %for fast
processing. Compressed formats are usually adopted. Although %the files are
large, the data can be easily decomposed because the %processing is done on each
sequence (`string') at a time

@@

%In research there are no strict constraints on the time to complete %the
analysis.  In patient care, time could determine life, death or %money, but we
have no concrete case yet.
%Quality in this case refers to the accuracy of the results %obtained. This is
essential for research and patient care.

@@

%\item Most of the dynamism is in infrastructure, not in the application
(e.g. use of PilotJob)
%\item Application dynamism: location of file (e.g. if file turns out to be not
local, look up in registry-- but not do this again)

@@

%A physics group is interested in a specific physics channel (underlying
physical process) as recorded by the Atlas experiment at the LHC (Large Hadron
Collider) at CERN in Geneva. The set of available data grows steadily over time
as more data are collected and reconstructed.

@@

% Both physics groups and individuals run jobs as they see fit, either reading
the reconstructed data or reading intermediate datasets created by themselves or
their colleagues. If, in the future, resources should prove inadequate then
adjustments to working practices to coordinate processing may be made to
increase physics output.
instruments, etc.)
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a
mixture of gLite software, experiment specific software and physics group
specific software. WLCG has 250,000 cores distributed over 140 sites and with
100~PB of disk. CERN is connected to the eleven `Tier 1' national sites by a
dedicated optical network. This may be extended in the future.

@@

%The data being processed can be considered as dynamic as it grows steadily as
new raw data are collected and the basic reconstruction program is run to
produce information about specific particles involved in a collision. This
reconstruction is rerun periodically (two or three times per year is expected)
as calibrations of the detector are improved. Reconstruction algorithms are also
improved as are the physics groups' selection codes. One output of the
reconstruction is `tag' data - summary information written to a database to
facilitate event selection. This is not yet used very much. Calibration data are
held on an RDBMS with Frontier/Squid front ends.

@@

% The bulk of the data is individual event data at different stages of
refinement. This is held in a in-house representation of C++ serialized
objects. Though this format is not standard it is used by all LHC experiments.

@@

% Roughly 20~TB of data are collected every day. Multiple generations of
processed data are kept. Many data files are replicated on multiple sites; this
is done both by copying it to where it is likely to be useful and partly
dynamically as the need arises.

@@

%A copy of the file is made to the SE (Storage Element) at the site where the
job will run. From there it is either copied to a local disk or read from the
SE.

@@

%If some data are missed during processing the only effect is a reduction of
statistical precision in the final analysis. Accurate book keeping is important
to know which data have been processed. If the selection code has been improved
then naturally the physics groups would like to see the results as soon as
possible.

@@

%A physics group is interested in a specific physics channel (underlying
physical process) as recorded by the Atlas experiment at the LHC (Large Hadron
Collider) at CERN in Geneva. The set of available data grows steadily over time
as more data are collected and reconstructed. It would not be efficient to allow
individual physics groups to process this large volume of data in an
uncoordinated manner as it requires reading all available data.

@@

%Physics groups register their analysis code to be run on the next scheduled
analysis ``train''. The train leaves every two weeks and reads through all the
data, runs the physics selection code of each group and writes out the selected
events and output from the computation to produce relatively small data sets for
each physics group. The groups then use these selections to perform more ad-hoc
analysis.
instruments, etc.)
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a
mixture of gLite software, experiment specific software and physics group
specific software. WLCG has 250,000 cores distributed over 140 sites and with
100~PB of disk. There is no wide area dedicated network infrastructure.

@@

%The data being processed can be considered as dynamic as it grows steadily as
new raw data are collected and the basic reconstruction program is run to
produce information about specific particles involved in a collision. This
reconstruction is rerun periodically (two or three times per year is expected)
as calibrations of the detector are improved. Reconstruction algorithms are also
improved as are the physics groups' selection code. Calibration data are held on
an RDBMS with Frontier/Squid front ends.

@@

% The bulk of the data is individual event data at different stages of
refinement. The is held in a non-standard representation of C++ serialized
objects.

@@

% Currently the reconstructed data (excluding copies) is about m TB and is
growing at about n TB per month. Many data files are replicated on multiple
sites. This is done both by copying it to where it is likely to be useful and
partly dynamically as the need arises.

@@

%This is mostly by copying the file to where the job will run - though some
remote file access is also used.

@@

%If some data are missed during processing the only effect is a reduction of
statistical precision in the final analysis. If the selection code has been
improved then naturally the physics groups would like to see the results as soon
as possible.

@@

%\item Analysis of data to generate event streams $\rightarrow$ lead to the %
positioning of telescopes

@@

%activities, despite the fact that they are different in many respects, %and
I'll try and distinguish between them where appropriate.
%Let me know if you want clarification or further information - in some cases I
wasn't quite sure what you were after.

@@

%These are now typically implemented in relational databases, with SQL
interfaces %implemented through web pages, and the principal goal of the Virtual
Observatory %(VO) is to standardize access to these, and other astronomical data
resources, %so that astronomers can readily perform multi-wavelength analyses,
combining all %the extant data on particular sources.

@@

% 3. What infrastructure is used? (including compute, data, network, %
instruments, etc.)

@@

%surveys---are typically implemented in relational databases running on
high-spec servers %or clusters thereof. The compute requirements of the VO are
currently modest, %but are likely to increase over time. Astronomers typically
use normal academic %networks, but more specialized networking is required in
some cases: e.g., sky %survey data is moved from Cambridge to Edinburgh using
the UKLight project's %dedicated optical network, and radio interferometers
require high-bandwidth %fibre connections from each antenna to the correlator.
% 4. What dynamic data is used in the application?  %(a) What are the types of
data, (b) What is the size of the data set(s)?

@@

%trigger follow-up observations on other telescopes. These are small XML
messages %containing basic information about the position of a transient source
and a %small amount of other metadata which the operators of a given telescope
can use %to define triggers for follow-up observations of interest to them or
their %community.

@@

%The time constraints differ for different types of transient source, but in
some %cases notification must be made within a few minutes if it is to be
useful; the %overheads associated with setting up a follow-up observation mean
that much %shorter time scales are practically irrelevant, although they are
still %scientifically interesting. The quality constraints centre on providing a

@@

%transient will be some sort of noise event, so fairly sophisticated filtering
is %required to bring the false positive rate down to something acceptable,
given %the cost of observing time on cutting edge telescopes.

@@

%\alnote{Is the workflow part still part of the app as defined before with the
stages: data transformation \& decision and storage?} \katznote{I think this is
data transformation.}\alnote{ok}

@@

Pan-STARRS, a contemporary sky survey with similar goals as LSST, uses a
scripted programming model on a cluster for image processing of telescope frames
to extract object attributes as a CSV file.  This is followed by followed by
scientific workflows, based on Trident, that load the files into individual
databases daily, and merges them with a master distributed-database on a weekly
basis on an HPC cluster~\cite{Simmhan:Building:2009}. Data extraction and mining
operations by astronomers are composed using a pseudo-SQL language with user
defined functions and submitted for execution on the master database using a
batch system.

@@

%\item What is the science/engineering problem? What is the purpose of the
application?

@@

%Processing of data (data analysis and detection of new sources) from each image
should be done while the next image is being taken by the telescope.  This means
that if anything unusual is detected, normal observation can be
interrupted. Other observing resources can also be notified instantly, so they
can observe the same event. As data are collected, they will be added to all the
data previously detected from the same location of sky to create a very deep
master image. LSST will also build up a database of all known moving objects.
%Every time a new image of the sky is obtained, the master image will be
subtracted from it. The result is a subtracted image that contains the
difference between the current sky image and its average state.
%This subtracted image is then processed by a cluster of computers with the
following steps:

@@

%\item Query object catalog (knows orbits of all known objects) to find objects
which are expected to appear in subtracted image, given the area of sky, time of
day.
%\item Cross-match expected objects with sources in subtracted image. The result
is the unmatched-source catalog, which contains sources that can't be matched
with a previously known object
%\item Use unmatched-source catalogue to compute pairs of detections separated
by short time intervals, called tracklets. A tracklet is an observable, short
section of orbit. In order to create pairs of detections, all objects in the
unmatched-source catalog are queried against the orbit catalog, in an attempt to
obtain data about these objects from earlier observations. If earlier
observations and current observations can be linked to form tracklets, they are
stored in the tracklet catalog.
%\item Use tracklet catalog to attempt to link these tracklets over a larger
time window. This is achieved by again querying the orbit catalog, this time
looking for observations of the same objects even further back in time. If
matches can be found and the tracklets can be extrapolated out into longer
sections of orbit, then they are added as new orbits to the orbit catalog.
%\item Update orbit catalog with redetections of known objects. Each rediscovery
provides information that improves the known orbits.
%\item Attempt to classify all entries in the orbit catalog (i.e., all the
objects which now have known orbits). If any Near Earth Objects are detected to
be passing close to the earth, alerts are generated to astronomers so that
follow up observations can be scheduled.
%\item Compare with other observatories: If an unknown object is found that
cannot be classified locally

@@

%\item local observatory chooses from accepting observatories one to obtain data
from

@@

%\item local observatory decides what to do - classify the object, contact a
human, call for more observations (automatically to other observatories, that
decide whether or not to accept, and when)

@@

%large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)

@@

%LSST will generate 36~GB of data every 30 seconds, and over a 10-hour winter
night, will collect up to 30~TB.
%3 What infrastructure will be used? (including compute, data, network,
instruments, etc.)

@@

%Some will be analyzed at the mountain top, discoveries will be sent to other
centers.
%6 Are there time (or quality) constraints on the application?  (If so, what are
they?)

@@

%7 Please feel free to also talk about the current state of the application, if
it exists today, and any specific gaps that you know will need to be overcome.

@@

%The Climate Data Analysis Tools (CDAT)~\cite{CDAT} are a scriptable %(Python)
set of packages and tools that are currently heavily used for %analysis of
climate data.  CDAT has been developed by LLNL's Program %for Climate Model
Diagnosis and Intercomparison (PCMDI), and is often %used with data from the
Earth System Grid (ESG)~\cite{ESG}.  The most %common current usage of these
tools with the data is that a user %downloads a set of data from ESG to a local
resource, then uses the %CDAT to analyze the data.  However, the PCMDI group is
now planning to %support a new usage mode, where data transfer can be scripted.
%Initial development of this capability is planned for February 2011, %using the
Globus online data transfer service~\cite{globusOnline} as a %scriptable
component.

@@

%Aimed at supporting international CMIP/IPCC intercomparison activity.  %Three
stages:

@@

%\item Climate centers run prescribed set of common experiments; produce 2-10 PB
of data

@@

%\item data can also come from sensors - here, the data would still be
post-processed before it would be published.

@@

%\item Centers can publish their own output data or send it to another center to
publish data generated over \~{}2 year period (then post-processed and published
over a few more months)

@@

%\item ESGF develops and deploys a federated network of gateways and associated
data nodes
%\item As models run at each center, output data is post-processed into packages
with common formats and prescribed metadata conventions
%\item Most centers will deploy data node software stack, and using this, they
manage the data from their experiments

@@

% \item the data node software stack scans the data, makes sure it has the right
metadata fields, does QA/QC, build a set of catalogs of the prepared data

@@

% \item also can provide product services that abstract the dataset in other
ways - get whole file, subset, browse, etc.

@@

% \item When the center is happy with the data/catalog in the data node, they
publish it to a host/affiliated gateway

@@

% \item them the gateway shares this catalog with other gateways so that all
gateways have a consistent view of all the published data

@@

% \item replication: copies catalog to replica site, do data transfer to new
data node, then publish catalog, push catalog to (this) gateway associated with
new data node

@@

% \item track which users have downloaded data - if a problem is found in a data
set or if new data is added to a data set, notify users who have
downloaded/replicated it

@@

% \item Properties: data-intensive, distributed gateways and data nodes, data
appears in the system over time (dynamic)

@@

% \item browse/search a catalog at any gateway and locate data, which might be
hosted by a data node affiliated with another gateway. (can output a wget script
that can later fetch the data)

@@

% \item download data via http or GridFTP or access product services (uses data
retrieval syntax (DRS) - can be scripted)

@@

% \item abstractions that can be used by a programmer: data is stored in NetCDF
format, uses CF metadata conventions, can be accessed using DRS

@@

% \item Many applications are multi-model analyses - many users want to look at
the same parts of the output of some/all of the models, to understand if and how
the models differ
% \item Some centers will gather some/all of the core archive (plus more,
perhaps) on local systems for local users to perform ``power analysis''

@@

%\subsubsection{old material follows} %1. What is the science/engineering
problem? What is the purpose of the %application?
%With ITER scheduled to operate before 2020, plasma fusion has become %one of
the highest priority research items for DOE. Scalability of %first-principles
fusion codes is an important part of fusion research %needed to obtain
predictive capability of plasma performance with %first-principles physics
fidelity. We want to build a realistic %full-distribution kinetic code, in
realistic ITER geometry, combining %the strengths of each of the existing
gyrokinetic codes (GEM, GTS, %XGC, and GTC), and utilize the most up-to-date
computer science, %hardware design and applied mathematics technologies
developed. The %application that we will develop will be capable of 1)
discovering %Macroscopic Magneto Hydro Dynamics activity in burning plasmas, 2)
%Understand Energetic particle effects in fusion reactors, 3) %Understand the
heating in a fusion reactor, and 4) Understand the %interaction of the plasma at
the material boundary.

@@

% i) large (exascale) data and large (exascale) computing (e.g., data mining,
analytics, bioinformatics)

@@

%The applications we currently run (XGC1, XGCP, GTC, GTS) run on all of the
leadership class facilities, and require very large resources for the modest
problems which we are currently solving. In the next generation (exascale) we
will not only need all of the computing power that will be available at the time
to understand ITER, we will also need to understand all of the data that will be
generated during the simulation. Our techniques have been to examine and
understand the data in situ, to reduce the total amount of data produced in the
simulation. Today we have produced O(100 TB/week) for a simulation using our
ADIOS middleware incorporated in our simulations. Using our staging techniques,
we believe that we will only produce O(10 PB) on an exascale machine. (I don't
know if this is considered modest dat, or extreme size data)?

@@

%1 PB of data -- 10 PB of data, for an exascale size computation run in O(100
hours).  %Today we achieve about 6\% of peak with excellent scalability to 220K
%cores on the Cray XT systems. It is %difficult to predict how we will run on

@@

%3. What infrastructure will be used? (including compute, data, %network,
instruments, etc.)

>>>>

\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{enumerate}
\usepackage{longtable}
%\usepackage{authblk} % Is used internally within the CPE Class
%\usepackage[firstpage]{draftwatermark}

@@

%\item Compute resource selection based on data location, processing profile and/or network capacity

@@

%\item Collection and processing of monitoring information to steer grid enactment.

@@

%\item Infrastructure monitoring for the execution of workflows to
%  manage failure
%\item the application is sequence alignment (split data (this is where there is dynamic decision making) to execute the alignment in parallel)
%\item DTI Imaging: two cases are interesting: DTI atlas (split, run, merge, split, merge - reduction from 10000 tasks to 1 output); PCA for classification of patients/control

@@

%The application performs analysis of DNA sequencing data for various
%goals in life science research, e.g., Mutation screening, Virus
%discovery, Genome-wide Associations (GWA), Linkage analysis,
%Comparison of bacterial genomes, microRNA expression and Exome
%sequencing. The data analysis is composed of various methods that are
%implemented by a bioinformatician, whereas the data is typically owned
%by the life scientist.

@@

%bioinformatician, excluding steps related to data acquisition
%(transformation of raw images into sequences of amino acids) and
%posterior biostatistics and epidemiological analysis.
%The data analysis is implemented as a pipeline of generic and
%customized components, mostly third-party public software tools. The
%pipeline may differ for each specific study, depending on the data
%acquisition and the research question. Typically it includes data
%preparation (file conversion, reformatting and filtering); alignment
%(comparison) with some reference database as the human genome;
%post-processing (output conversion, reformatting, statistical

@@

%computing-intensive step, which became prohibitive for sequential
%execution in next generation sequence experiments. All data is stored
%in files that are often completely read in memory for fast
%processing. Compressed formats are usually adopted. Although the files
%are large, the data can be easily decomposed because the processing is
%done on each sequence (`string') at a time.

@@

%Grid storage: Storage Elements of the Dutch e-Science Grid
%(www.biggrid.nl)

@@

%Between data server and grid storage: public network (XXX); lightpath
%(near future)
%Between grid storage and worker nodes: public (XXX) or site (XXX)
%network

@@

%Local Computing: clusters and servers at the AMC and workstations
%(laptops, desktops) anywhere Grid Computing: Dutch Grid
%(www.biggrid.nl), part of EGI
%Originally all steps were performed in some local server or on the
%user's workstation. The visualization is still performed in the user's
%(windows) workstation, but all the other steps are currently performed
%on the Dutch grid infrastructure. The pipeline is described as a grid
%workflow using the GWENDIA language
%(\url{http://dx.doi.org/10.1145/1645164.1645171}) and enacted on the

@@

%PS: although only the alignment step is computation-intensive, it is
%more practical to perform all the others on the grid to minimize data
%transport to/from the user's workstation, which is normally on low
%speed networks. That is, the computation is statically and manually
%taken to the data in this case. This can lead to sub-optimal resource
%usage.

@@

%Note: data is not really `dynamic' in this case -- the sequences are
%acquired once and analyzed many times. However the execution of this
%application on a distributed infrastructure has dynamic aspects when
%optimized resource usage is considered:

@@

%DNA sequences (strings) stored in open formats (FASTA, SFF, SAM, BAM)
%Bundles of results from alignment tools (custom format, binary?)

@@

%input data: Sequencing Data in the csFasta format, normally 25-35~GB;
%Quality files in the .qual format, normally 50-80~GB; Reference DB in
%the Fasta BS format: 3.2~GB (human genome) 140~MB (one chromosome)
%Intermediate files: Reference BWA index, normally 4.5GB (human genome)
%240~MB (one chromosome); Sequencing Data in the FastQ format
%(fastq.gz), normally 20-30~GB
%Output: Results in .sai format (direct output of BWA), normally
%2-3~GB; Results in .sam format, normally 55-75~GB; Results in .bam
%format, normally 20-30~GB

@@

%From a local file. The legacy code is wrapped to stage the inputs and
%outputs in/out the worker node.
%All data is stored in files that are often completely read in memory
%for fast processing. Compressed formats are usually adopted. Although
%the files are large, the data can be easily decomposed because the
%processing is done on each sequence (`string') at a time

@@

%In research there are no strict constraints on the time to complete
%the analysis.  In patient care, time could determine life, death or
%money, but we have no concrete case yet.
%Quality in this case refers to the accuracy of the results
%obtained. This is essential for research and patient care.

@@

%\item Most of the dynamism is in infrastructure, not in the application (e.g. use of PilotJob)
%\item Application dynamism: location of file (e.g. if file turns out to be not local, look up in registry-- but not do this again)

@@

%A physics group is interested in a specific physics channel (underlying physical process) as recorded by the Atlas experiment at the LHC (Large Hadron Collider) at CERN in Geneva. The set of available data grows steadily over time as more data are collected and reconstructed.

@@

% Both physics groups and individuals run jobs as they see fit, either reading the reconstructed data or reading intermediate datasets created by themselves or their colleagues. If, in the future, resources should prove inadequate then adjustments to working practices to coordinate processing may be made to increase physics output.
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a mixture of gLite software, experiment specific software and physics group specific software. WLCG has 250,000 cores distributed over 140 sites and with 100~PB of disk. CERN is connected to the eleven `Tier 1' national sites by a dedicated optical network. This may be extended in the future.

@@

%The data being processed can be considered as dynamic as it grows steadily as new raw data are collected and the basic reconstruction program is run to produce information about specific particles involved in a collision. This reconstruction is rerun periodically (two or three times per year is expected) as calibrations of the detector are improved. Reconstruction algorithms are also improved as are the physics groups' selection codes. One output of the reconstruction is `tag' data - summary information written to a database to facilitate event selection. This is not yet used very much. Calibration data are held on an RDBMS with Frontier/Squid front ends.

@@

%          The bulk of the data is individual event data at different stages of refinement. This is held in a in-house representation of C++ serialized objects. Though this format is not standard it is used by all LHC experiments.

@@

%          Roughly 20~TB of data are collected every day. Multiple generations of processed data are kept. Many data files are replicated on multiple sites; this is done both by copying it to where it is likely to be useful and partly dynamically as the need arises.

@@

%A copy of the file is made to the SE (Storage Element) at the site where the job will run. From there it is either copied to a local disk or read from the SE.

@@

%If some data are missed during processing the only effect is a reduction of statistical precision in the final analysis. Accurate book keeping is important to know which data have been processed. If the selection code has been improved then naturally the physics groups would like to see the results as soon as possible.

@@

%A physics group is interested in a specific physics channel (underlying physical process) as recorded by the Atlas experiment at the LHC (Large Hadron Collider) at CERN in Geneva. The set of available data grows steadily over time as more data are collected and reconstructed. It would not be efficient to allow individual physics groups to process this large volume of data in an uncoordinated manner as it requires reading all available data.

@@

%Physics groups register their analysis code to be run on the next scheduled analysis ``train''. The train leaves every two weeks and reads through all the data, runs the physics selection code of each group and writes out the selected events and output from the computation to produce relatively small data sets for each physics group. The groups then use these selections to perform more ad-hoc analysis.
%The WLCG (Worldwide LHC (Large Hadron Collider) Computing Grid) is used with a mixture of gLite software, experiment specific software and physics group specific software. WLCG has 250,000 cores distributed over 140 sites and with 100~PB of disk. There is no wide area dedicated network infrastructure.

@@

%The data being processed can be considered as dynamic as it grows steadily as new raw data are collected and the basic reconstruction program is run to produce information about specific particles involved in a collision. This reconstruction is rerun periodically (two or three times per year is expected) as calibrations of the detector are improved. Reconstruction algorithms are also improved as are the physics groups' selection code. Calibration data are held on an RDBMS with Frontier/Squid front ends.

@@

%          The bulk of the data is individual event data at different stages of refinement. The is held in a non-standard representation of C++ serialized objects.

@@

%          Currently the reconstructed data (excluding copies) is about m TB and is growing at about  n TB per month. Many data files are replicated on multiple sites. This is done both by copying it to where it is likely to be useful and partly dynamically as the need arises.

@@

%This is mostly by copying the file to where the job will run - though some remote file access is also used.

@@

%If some data are missed during processing the only effect is a reduction of statistical precision in the final analysis. If the selection code has been improved then naturally the physics groups would like to see the results as soon as possible.

@@

%\item Analysis of data to generate event streams $\rightarrow$ lead to the
%  positioning of telescopes

@@

%activities, despite the fact that they are different in many respects,
%and I'll try and distinguish between them where appropriate.
%Let me know if you want clarification or further information - in some cases I wasn't quite sure what you were after.

@@

%These are now typically implemented in relational databases, with SQL interfaces
%implemented through web pages, and the principal goal of the Virtual Observatory
%(VO) is to standardize access to these, and other astronomical data resources,
%so that astronomers can readily perform multi-wavelength analyses, combining all
%the extant data on particular sources.

@@

%      instruments, etc.)

@@

%surveys---are typically implemented in relational databases running on high-spec servers
%or clusters thereof. The compute requirements of the VO are currently modest,
%but are likely to increase over time. Astronomers typically use normal academic
%networks, but more specialized networking is required in some cases: e.g., sky
%survey data is moved from Cambridge to Edinburgh using the UKLight project's
%dedicated optical network, and radio interferometers require high-bandwidth
%fibre connections from each antenna to the correlator.
%(a) What are the types of data, (b) What is the size of the data set(s)?

@@

%trigger follow-up observations on other telescopes. These are small XML messages
%containing basic information about the position of a transient source and a
%small amount of other metadata which the operators of a given telescope can use
%to define triggers for follow-up observations of interest to them or their
%community.

@@

%The time constraints differ for different types of transient source, but in some
%cases notification must be made within a few minutes if it is to be useful; the
%overheads associated with setting up a follow-up observation mean that much
%shorter time scales are practically irrelevant, although they are still
%scientifically interesting. The quality constraints centre on providing a

@@

%transient will be some sort of noise event, so fairly sophisticated filtering is
%required to bring the false positive rate down to something acceptable, given
%the cost of observing time on cutting edge telescopes.

@@

@@

Pan-STARRS, a contemporary sky survey with similar goals as LSST,  uses a scripted programming model
on a cluster for image processing of telescope frames to extract object attributes as a CSV
file.
%\alnote{Is the workflow part still part of the app as defined before with the stages: data transformation \& decision and storage?} \katznote{I think this is data transformation.}\alnote{ok}
This is followed by followed by scientific workflows, based on Trident, that load the files
into individual databases daily, and merges them with a master distributed-database on a weekly basis on an
HPC cluster~\cite{Simmhan:Building:2009}. Data extraction and mining operations by astronomers are composed using a pseudo-SQL
language with user defined functions and submitted for execution on the master database using a batch system.

@@

%\item What is the science/engineering problem? What is the purpose of the application?

@@

%Processing of data (data analysis and detection of new sources) from each image should be done while the next image is being taken by the telescope.  This means that if anything unusual is detected, normal observation can be interrupted. Other observing resources can also be notified instantly, so they can observe the same event. As data are collected, they will be added to all the data previously detected from the same location of sky to create a very deep master image. LSST will also build up a database of all known moving objects.
%Every time a new image of the sky is obtained, the master image will be subtracted from it. The result is a subtracted image that contains the difference between the current sky image and its average state.
%This subtracted image is then processed by a cluster of computers with the following steps:

@@

%\item Query object catalog (knows orbits of all known objects) to find objects which are expected to appear in subtracted image, given the area of sky, time of day.
%\item Cross-match expected objects with sources in subtracted image. The result is the unmatched-source catalog, which contains sources that can't be matched with a previously known object
%\item Use unmatched-source catalogue to compute pairs of detections separated by short time intervals, called tracklets. A tracklet is an observable, short section of orbit. In order to create pairs of detections, all objects in the unmatched-source catalog are queried against the orbit catalog, in an attempt to obtain data about these objects from earlier observations. If earlier observations and current observations can be linked to form tracklets, they are stored in the tracklet catalog.
%\item Use tracklet catalog to attempt to link these tracklets over a larger time window. This is achieved by again querying the orbit catalog, this time looking for observations of the same objects even further back in time. If matches can be found and the tracklets can be extrapolated out into longer sections of orbit, then they are added as new orbits to the orbit catalog.
%\item Update orbit catalog with redetections of known objects. Each rediscovery provides information that improves the known orbits.
%\item Attempt to classify all entries in the orbit catalog (i.e., all the objects which now have known orbits). If any Near Earth Objects are detected to be passing close to the earth, alerts are generated to astronomers so that follow up observations can be scheduled.
%\item Compare with other observatories: If an unknown object is found that cannot be classified locally

@@

%\item local observatory chooses from accepting observatories one to obtain data from

@@

%\item local observatory decides what to do - classify the object, contact a human, call for more observations (automatically to other observatories, that decide whether or not to accept, and when)

@@

%large (exascale) data and large (exascale) computing (e.g., data mining, analytics, bioinformatics)

@@

%LSST will generate 36~GB of data every 30 seconds, and over a 10-hour winter night, will collect up to 30~TB.
%3 What infrastructure will be used? (including compute, data, network, instruments, etc.)

@@

%Some will be analyzed at the mountain top, discoveries will be sent to other centers.
%6 Are there time (or quality) constraints on the application?  (If so, what are they?)

@@

%7 Please feel free to also talk about the current state of the application, if it exists today, and any specific gaps that you know will need to be overcome.

@@

%The Climate Data Analysis Tools (CDAT)~\cite{CDAT} are a scriptable
%(Python) set of packages and tools that are currently heavily used for
%analysis of climate data.  CDAT has been developed by LLNL's Program
%for Climate Model Diagnosis and Intercomparison (PCMDI), and is often
%used with data from the Earth System Grid (ESG)~\cite{ESG}.  The most
%common current usage of these tools with the data is that a user
%downloads a set of data from ESG to a local resource, then uses the
%CDAT to analyze the data.  However, the PCMDI group is now planning to
%support a new usage mode, where data transfer can be scripted.
%Initial development of this capability is planned for February 2011,
%using the Globus online data transfer service~\cite{globusOnline} as a
%scriptable component.

@@

%Aimed at supporting international CMIP/IPCC intercomparison activity.
%Three stages:

@@

%\item Climate centers run prescribed set of common experiments; produce 2-10 PB of data

@@

%\item data can also come from sensors - here, the data would still be post-processed before it would be published.

@@

%\item Centers can publish their own output data or send it to another center to publish data generated over \~{}2 year period (then post-processed and published over a few more months)

@@

%\item ESGF develops and deploys a federated network of gateways and associated data nodes
%\item As models run at each center, output data is post-processed into packages with common formats and prescribed metadata conventions
%\item Most centers will deploy data node software stack, and using this, they manage the data from their experiments

@@

%  \item the data node software stack scans the data, makes sure it has the right metadata fields, does QA/QC, build a set of catalogs of the prepared data

@@

%                \item also can provide product services that abstract the dataset in other ways - get whole file, subset, browse, etc.

@@

%    \item When the center is happy with the data/catalog in the data node, they publish it to a host/affiliated gateway

@@

%          \item  them the gateway shares this catalog with other gateways so that all gateways have a consistent view of all the published data

@@

%          \item  replication: copies catalog to replica site, do data transfer to new data node, then publish  catalog, push catalog to (this) gateway associated with new data node

@@

%          \item  track which users have downloaded data - if a problem is found in a data set or if new data is added to a data set, notify users who have downloaded/replicated it

@@

%    \item  Properties: data-intensive, distributed gateways and data nodes, data appears in the system over time (dynamic)

@@

%         \item  browse/search a catalog at any gateway and locate data, which might be hosted by a data node affiliated with another gateway. (can output a wget script that can later fetch the data)

@@

%        \item  download data via http or GridFTP or access product services (uses data retrieval syntax (DRS) - can be scripted)

@@

%    \item  abstractions that can be used by a programmer: data is stored in NetCDF format, uses CF metadata conventions, can be accessed using DRS

@@

%    \item  Many applications are multi-model analyses - many users want to look at the same parts of the output of some/all of the models, to understand if and how the models differ
%    \item  Some centers will gather some/all of the core archive (plus more, perhaps) on local systems for local users to perform ``power analysis''

@@

%1. What is the science/engineering problem? What is the purpose of the
%application?
%With ITER scheduled to operate before 2020, plasma fusion has become
%one of the highest priority research items for DOE. Scalability of
%first-principles fusion codes is an important part of fusion research
%needed to obtain predictive capability of plasma performance with
%first-principles physics fidelity. We want to build a realistic
%full-distribution kinetic code, in realistic ITER geometry, combining
%the strengths of each of the existing gyrokinetic codes (GEM, GTS,
%XGC, and GTC), and utilize the most up-to-date computer science,
%hardware design and applied mathematics technologies developed. The
%application that we will develop will be capable of 1) discovering
%Macroscopic Magneto Hydro Dynamics activity in burning plasmas, 2)
%Understand Energetic particle effects in fusion reactors, 3)
%Understand the heating in a fusion reactor, and 4) Understand the
%interaction of the plasma at the material boundary.

@@

%        i) large (exascale) data and large (exascale) computing (e.g., data mining, analytics, bioinformatics)

@@

%The applications we currently run (XGC1, XGCP, GTC, GTS) run on all of the leadership class facilities, and require very large resources for the modest problems which we are currently solving. In the next generation (exascale) we will not only need all of the computing power that will be available at the time to understand ITER, we will also need to understand all of the data that will be generated during the simulation. Our techniques have been to examine and understand the data in situ, to reduce the total amount of data produced in the simulation. Today we have produced O(100 TB/week) for a simulation using our ADIOS middleware incorporated in our simulations. Using our staging techniques, we believe that we will only produce O(10 PB) on an exascale machine. (I don't know if this is considered modest dat, or extreme size data)?

@@

%1 PB of data -- 10 PB of data, for an exascale size computation run in O(100 hours).
%Today we achieve about 6\% of peak with excellent scalability to 220K
%cores on the Cray XT systems. It is
%difficult to predict how we will run on

@@

%3. What infrastructure will be used? (including compute, data,
%network, instruments, etc.)
%For Validation, we will need to compare our data from the computer to
